### 2021-11-30 chapter5/chapter.tex
.

\section{Function approximation/reconstruction}
We start our investigation by looking at how well the methods perform at approximating an initial condition. For the discontinuous Galerkin method, this means projecting the solution on to the set of polynomial functions which define the scheme, then reconstructing the solution. For the spatially-coherent uniformisation scheme, this means looking at the sub-cell averages. For the QBD-RAP scheme, this means computing the intial vector for the approximation, then reconstructing the solutions as described in Section~\ref{sec: closing}. First we investigate three closing vectors we can use for the reconstruction for the QBD-RAP, from which we find that the closing operator in (\ref{eqn:5462}) performs reasonably compared to the others, and we are able to theoretically prove that it converges. Therefore we use this reconstruction method throughout the rest of the chapter.

For this section we cosider approximating the initial condition over a single cell of width \(\Delta = 1\). To numerically evaluate intergals arising in the approximation step (inner products and cell averages) we use a simple trapezoidal rule with 10,001 function evaluations. Similarly, we use 10,001 function evaluations to approximate \(L^2\) norms, and also as a finite set of points overwhich to compute the KS statistics.

@@

First consider approximating the initial condition with density function \(1(x<0.5)\). In the left-hand side of Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} we plot the Kolmogorov-Smirnov error between the true CDF and the CDF constructed via the QBD-RAP approximation. Interestingly, in this case, the unnormalised error outperforms the other two reconstructions at low orders. It just so happens that, in this case, the error due to truncation of the unnormalised reconstruction destructively interfers with other errors in the reconstruction to cause the error to be lower at low orders for the unnormalised scheme. Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} also shows that the naive normalised and normalised reconstructions perform very similarly -- this is the case throughout much of this subsection. The difference between the naive normalised and normalised reconstructions is how they treat mass in the tail of a matrix exponential distribution (from \(2\Delta\) onward). Intuitively, there should be very little mass in the tail of the distribution as we are using conentrated matrix exponential distributions in the reconstruction.

@@

Figure~\ref{fig: pdf reconstructed} also show that both closing operators do not approximate the initial condition well at the right-hand side of the interval. This is a common feature, and drawback, of the QBD-RAP method. Perhaps there is a different reconstruction method or yet another closing operator which could aleviate this issue. 
\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{chapter5/figs/qbdrap_closing_vec/fun4/pdfs_formatted.pdf}
	\caption{Reconstructed PDFs using the unnormalised closing operator (blue solid line), normalised closing operator (green dash-dotted line), for various orders, and the true PDF which is \(f(x)=1\) (orange dashed line).}
	\label{fig: pdf reconstructed}
\end{figure}

Next we consider approximating the initial distribution with density \(-6x^2+6x\). Observing the left-hand panel of Figure~\ref{fig: fun 6 ks error qbdrap closing vecs}, which plots the KS error against the order of the reconstruction for the three closing operators, we once again see that the reconstruction using the unnormalised closing opeartor performs the worst, while the performance of the two normalised reconstructions is indistinguishable. However, if we instead look at the right-hand panel of Figure~\ref{fig: fun 6 ks error qbdrap closing vecs}, which plots the \(L^2\) error between the reconstructed PDF and the true PDF, we see that with this metric, then unnormalised closing operator performs better than the two normalised ones for orders 5 and above. 
\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/qbdrap_closing_vec/fun6/ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/qbdrap_closing_vec/fun6/l2_pdf_error_formatted.pdf}
	\caption{KS error between the true CDF, \(-2x^3+3x^2\), and the approximation (left) and \(L^2\) error between the true PDF \(-6x^2+6x\) and the approximation (right), for the three closing vectors considered; unnormalised (blue solid line with crosses), naive normalised (orange dashed line with diamonds) and normalised (green dash-dotted line with circles). The naive normalised (orange) an normalised closing vectors are coincident.}
	\label{fig: fun 6 ks error qbdrap closing vecs}
\end{figure}

The fact that the unnormalised closing operator outperforms the two normalised ones can be explain, once again, by the fact that, for the normalised operators we have `folded' the ME density function back on itself. In Figure~\ref{fig: pdf reconstructed quadratic} we plot the unnormalised and normalised reconstructions along with the true PDF, \(-6x^2+6x\). In Figure~\ref{fig: pdf reconstructed quadratic} we observe that, at the left of the plot, the normalised reconstructions over estimate the density function in this region, where as the unnormalised reconstruction looks to be doing better. Recall that the `folding' in the normalised closing operatot manifest itself as extra mass at the left of these plots compared to the unnormalised reconstructions. 
\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{chapter5/figs/qbdrap_closing_vec/fun6/pdfs_formatted.pdf}
	\caption{Reconstructed PDFs using the unnormalised closing operator (blue solid line), normalised closing operator (green dash-dotted line), for various orders, and the true PDF which is \(f(x)=-6x^2+6x\) (orange dashed line).}
	\label{fig: pdf reconstructed quadratic}
\end{figure}

Lastly, we consider approximating the intial distribution with PDF \(3e^{-3x}/(1-e^{-3})\). This density function is at a maximum at the left of the region. Considering what we have learnt so far about the unnormalised operator underestimating in this region, we expect that the unnormalised closing operator will perform relatively poorly in this case. Indeed, observing Figures~\ref{fig: fun 7 ks error qbdrap closing vecs} we see that the normalised reconstructions perform relatively well compared to the unnormalised reconstruction as measured by both error metrics (KS statistic and \(L^2\) norm on the PDF). If we observed plots of the PDFs (omitted), we would once again see that this is due to the loss of mass at the left-hand side of the region due to the truncation. 
\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/qbdrap_closing_vec/fun7/ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/qbdrap_closing_vec/fun7/l2_pdf_error_formatted.pdf}
	\caption{KS error between the true CDF, \((1-e^{-3x})/(1-e^{-3})\), and the approximation (left) and \(L^2\) error between the true PDF and the approximation for the three closing vectors considered; unnormalised (blue solid line with crosses), naive normalised (orange dashed line with diamonds) and normalised (green dash-dotted line with circles). The naive normalised (orange) an normalised closing vectors are coincident.}
	\label{fig: fun 7 ks error qbdrap closing vecs}
\end{figure}

>>>>

\section{Function approximation/reconstruction}
We start our investigation by looking at how well the methods perform at approximating an initial condition. For the discontinuous Galerkin method, this means projecting the solution on to the set of polynomial functions which define the scheme, then reconstructing the solution. For the spatially-coherent uniformisation scheme, this means looking at the sub-cell averages. For the QBD-RAP scheme, this means computing the initial vector for the approximation, then reconstructing the solutions as described in Section~\ref{sec: closing}. First we investigate three closing vectors we can use for the reconstruction for the QBD-RAP, from which we find that the closing operator in (\ref{eqn:5462}) performs reasonably compared to the others, and we are able to theoretically prove that it converges. Therefore we use this reconstruction method throughout the rest of the chapter.

For this section we consider approximating the initial condition over a single cell of width \(\Delta = 1\). To numerically evaluate integrals arising in the approximation step (inner products and cell averages) we use a simple trapezoidal rule with 10,001 function evaluations. Similarly, we use 10,001 function evaluations to approximate \(L^2\) norms, and also as a finite set of points over which to compute the KS statistics.

@@

First consider approximating the initial condition with density function \(1(x<0.5)\). In the left-hand side of Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} we plot the Kolmogorov-Smirnov error between the true CDF and the CDF constructed via the QBD-RAP approximation. Interestingly, in this case, the unnormalised error outperforms the other two reconstructions at low orders. It just so happens that, in this case, the error due to truncation of the unnormalised reconstruction destructively interferes with other errors in the reconstruction to cause the error to be lower at low orders for the unnormalised scheme. Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} also shows that the naive normalised and normalised reconstructions perform very similarly -- this is the case throughout much of this subsection. The difference between the naive normalised and normalised reconstructions is how they treat mass in the tail of a matrix exponential distribution (from \(2\Delta\) onward). Intuitively, there should be very little mass in the tail of the distribution as we are using concentrated matrix exponential distributions in the reconstruction.

@@

Figure~\ref{fig: pdf reconstructed} also show that both closing operators do not approximate the initial condition well at the right-hand side of the interval. This is a common feature, and drawback, of the QBD-RAP method. Perhaps there is a different reconstruction method or yet another closing operator which could alleviate this issue. 
\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{chapter5/figs/qbdrap_closing_vec/fun4/pdfs_formatted.pdf}
	\caption{Reconstructed PDFs using the unnormalised closing operator (blue solid line), normalised closing operator (green dash-dotted line), for various orders, and the true PDF which is \(f(x)=1\) (orange dashed line).}
	\label{fig: pdf reconstructed}
\end{figure}

Next we consider approximating the initial distribution with density \(-6x^2+6x\). Observing the left-hand panel of Figure~\ref{fig: fun 6 ks error qbdrap closing vecs}, which plots the KS error against the order of the reconstruction for the three closing operators, we once again see that the reconstruction using the unnormalised closing operator performs the worst, while the performance of the two normalised reconstructions is indistinguishable. However, if we instead look at the right-hand panel of Figure~\ref{fig: fun 6 ks error qbdrap closing vecs}, which plots the \(L^2\) error between the reconstructed PDF and the true PDF, we see that with this metric, then unnormalised closing operator performs better than the two normalised ones for orders 5 and above. 
\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/qbdrap_closing_vec/fun6/ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/qbdrap_closing_vec/fun6/l2_pdf_error_formatted.pdf}
	\caption{KS error between the true CDF, \(-2x^3+3x^2\), and the approximation (left) and \(L^2\) error between the true PDF \(-6x^2+6x\) and the approximation (right), for the three closing vectors considered; unnormalised (blue solid line with crosses), naive normalised (orange dashed line with diamonds) and normalised (green dash-dotted line with circles). The naive normalised (orange) an normalised closing vectors are coincident.}
	\label{fig: fun 6 ks error qbdrap closing vecs}
\end{figure}

The fact that the unnormalised closing operator outperforms the two normalised ones can be explain, once again, by the fact that, for the normalised operators we have `folded' the ME density function back on itself. In Figure~\ref{fig: pdf reconstructed quadratic} we plot the unnormalised and normalised reconstructions along with the true PDF, \(-6x^2+6x\). In Figure~\ref{fig: pdf reconstructed quadratic} we observe that, at the left of the plot, the normalised reconstructions over estimate the density function in this region, where as the unnormalised reconstruction looks to be doing better. Recall that the `folding' in the normalised closing operator manifest itself as extra mass at the left of these plots compared to the unnormalised reconstructions. 
\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{chapter5/figs/qbdrap_closing_vec/fun6/pdfs_formatted.pdf}
	\caption{Reconstructed PDFs using the unnormalised closing operator (blue solid line), normalised closing operator (green dash-dotted line), for various orders, and the true PDF which is \(f(x)=-6x^2+6x\) (orange dashed line).}
	\label{fig: pdf reconstructed quadratic}
\end{figure}

Lastly, we consider approximating the initial distribution with PDF \(3e^{-3x}/(1-e^{-3})\). This density function is at a maximum at the left of the region. Considering what we have learnt so far about the unnormalised operator underestimating in this region, we expect that the unnormalised closing operator will perform relatively poorly in this case. Indeed, observing Figures~\ref{fig: fun 7 ks error qbdrap closing vecs} we see that the normalised reconstructions perform relatively well compared to the unnormalised reconstruction as measured by both error metrics (KS statistic and \(L^2\) norm on the PDF). If we observed plots of the PDFs (omitted), we would once again see that this is due to the loss of mass at the left-hand side of the region due to the truncation. 
\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/qbdrap_closing_vec/fun7/ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/qbdrap_closing_vec/fun7/l2_pdf_error_formatted.pdf}
	\caption{KS error between the true CDF, \((1-e^{-3x})/(1-e^{-3})\), and the approximation (left) and \(L^2\) error between the true PDF and the approximation for the three closing vectors considered; unnormalised (blue solid line with crosses), naive normalised (orange dashed line with diamonds) and normalised (green dash-dotted line with circles). The naive normalised (orange) an normalised closing vectors are coincident.}
	\label{fig: fun 7 ks error qbdrap closing vecs}
\end{figure}

\subsection{Comparison of methods}
Here we compare the ability of the QBD-RAP, uniformisation, and DG methods to reconstruct initial conditions.

\section{Travelling wave}
Here we investigate the performance the different schemes for approximating transient distributions of one-dimensional travelling wave problems with varying initial conditions. Consider a (trivial) fluid queue with one phase, generator \(\bs T=[0]\) and rate \(c=1\). The PDE (if/when it exists) which describes this system is 
\[\cfrac{\partial}{\partial t} f(x,t) = \cfrac{\partial}{\partial x} f(x,t),\]
where \(f(x,t)\) is the density at time \(t\). Given an initial condition, \(f(x,0)\), solutions to this problem are given by 
\[f(x,t) = f(x-t,0)\]
so the solution at time \(t\) is just a shift in the initial condition \(t\) units to the right. We suppose that the fluid queue is bounded, with a lower boundary \(x=0\) and upper boundary \(x=10\).

We use the QBD-RAP, uniformisation and DG schemes to discretise the solution in space. We discretise the interval \([0,10]\) into \(10\) cell, each of width 1. We then
